\LARGE \bf Multi-Armed Bandit Models for 2D Grasp Planning with Uncertainty

### 2015-06-15 paper.tex
.

\begin{abstract}
For applications such as warehouse order fulfillment, robot grasps must be robust to uncertainty arising from sensing,
mechanics, and control.  One way to achieve robustness is to evaluate the performance of candidate grasps by sampling
perturbations in shape, pose, and gripper approach and to compute the probability of force closure for each candidate to
identify a grasp with the highest expected quality. Since evaluating the
quality of each grasp is computationally demanding, prior work has turned to cloud computing. To improve computational efficiency and to extend this work, we consider how
Multi-Armed Bandit (MAB) models for optimizing decisions can be applied in this context. We formulate robust grasp
planning as a MAB problem and evaluate convergence times towards an optimal grasp candidate using 100 objects shapes from
the Brown Vision 2D Lab Dataset with 1000 grasp candidates per object. We consider the case where shape uncertainty is represented as a Gaussian process implicit surface (GPIS) with Gaussian uncertainty in pose, gripper approach angle, and coefficient of friction. 
We find that Thompson Sampling and the Gittins index MAB methods converged to within 3\% of the optimal grasp up to 10x faster than uniform allocation and 5x faster than iterative pruning. 
\end{abstract}

@@

    \caption{Number of samples versus the normalized probability of force closure $P_F$ for the best estimated grasp 
        after $t$ samples, $P_F(\Gamma_{\bar{k}, t})$, out of 1000 candidate grasps using uniform allocation, iterative pruning (eliminating candidates that perform poorly on initial samples), and our proposed Multi-Armed Bandit (MAB) algorithms (Gittins indices and Thompson Sampling).
    %We evaluated performance on 100 objects randomly selected from from the Brown Vision 2D Lab Dataset \cite{brown} with 1000 uniformly distributed candidate grasp plans.
    The normalized $P_F$ is given by the ratio of $P_F(\Gamma_{\bar{k}, t})$ to the highest $P_F$ value in the candidate
    grasp set $P_F(\Gamma^*)$ averaged over 100 independent runs on randomly selected objects from the Brown Vision 2D Dataset \cite{brown}.
    The highest quality grasp was determined by brute force search over all candidate grasps (which requires 10x more
    iterations than all methods shown)~\cite{kehoe2012toward}.
    Uniform allocation and iterative pruning converge to within 3$\%$ of the highest quality grasp (the dashed grey line) in approximately 40,000 and 20,000 iterations, respectively. In comparison, the MAB methods both converge in approximately 4,000 iterations. 
%    $P_F(\Gamma_{\bar{k}})$ (ratio of $P_F(\Gamma_{\bar{k}})$ to $P_F(\Gamma^*)$)
%    
%    Normalized grasp quality over time of the returned grasp for 4 sample-based planning methods.  Normalization divides the current quality of grasp, $P_F(\Gamma_{\bar{k}})$ returned by the highest probability of force closure in the candidate grasp set $P_F(\Gamma^*)$. The highest quality grasp was determined by brute force search over all candidate grasps (which required 10x  more time than any of these methods \cite{kehoe2012toward}).  Data is based on 100 objects randomly selected from from the Brown Vision 2D Lab Dataset \cite{brown} with 1000 uniformly distributed candidate grasp plans. Monte Carlo methods such as Uniform sampling and iterative pruning (eliminating candidates that perform poorly on initial samples) eventually converge to within 3$\%$ of the optimal (the dashed grey line), requiring approximately 40,000 and 20,000 iterations.  Two Multi-Arm Bandit methods perform significantly better, converging after only 4000 iterations..
	}
    \label{fig:grasp_quality}
\end{figure}

@@

We compare Thompson sampling and Gittins indices, two popular algorithms for solving the MAB problem,
against uniform allocation and an adaptive sampling method known as iterative pruning, which iteratively reduces the set
of candidate grasps based on the sample mean~\cite{kehoe2012toward}. 
In terms of the number of samples needed to find a grasp with highest estimated probability of force closure on objects in the Brown Vision 2D Dataset, a dataset of 2D planar objects~\cite{brown, christopoulos2007handling}. Our simulation results show that Thompson Sampling, a MAB algorithm, required $5$x fewer samples than iterative
pruning and $10$x fewer samples than uniform allocation to determine a grasp within $3\%$ of the estimated highest probability of force closure
grasp among a set of 1000 grasp candidates per object and averaged over 100 objects.

@@

One benefit of the Beta prior on Bernoulli reward distributions is that updates to the belief distribution after observing rewards from arm pulls can be derived in closed form.
At timestep $t=0$, we pull arm $k$ and observe reward $R_{k,0}$, where $R_{k,0} \in \lbrace 0, 1\rbrace$.
The posterior of the Beta after this observation is $\alpha_{k, 1} = \alpha_{k, 0} + R_{k,0}$,  $\beta_{k, 1} = \beta_{k, 0} + 1-R_{k,0}$, where $\alpha_{k,0}$ and $\beta_{k,0}$ are the prior shape parameters for arm $k$ before any samples are evaluated.

@@

\subsubsection{Thompson Sampling}
The computational cost of the Gittins index method can increase exponentially in time as the discount factor approaches $1$.
However, in the case of finding the best arm, we want to plan for long term reward and thus want $\gamma$ as close to $1$ as possible.
Due to computational constraints we must use a smaller $\gamma$, but this can lead to the algorithm pulling only the most promising arm for many iterations~\cite{kelly1981multi}.

@@

To try and reduce the number of samples needed, we instead maximize the sum of $P_F$ values for each sampled grasp  $\Gamma_{k,t}$ at time $t$ up to a given time $T_s$:
\begin{align}
    \underset{\Gamma_{k,} \in \mG}{\text{max }} \sum \limits_{t=1}^
    {T_s} P_F \left( \Gamma_{k,t}\right)  \label{eq:problem_def}
\end{align}
\noindent This attempts to perform as well as Equation~\ref{eq:old_problem_def} in as few samples as possible~\cite{srinivas2009gaussian}. We then formulate the  problem as a MAB model and compare two different Bayesian MAB algorithms, Thompson sampling and Gittins indices.

>>>>

\begin{abstract}
For applications such as warehouse order fulfillment, robot grasps must be robust to uncertainty arising from sensing,
mechanics, and control.  One way to achieve robustness is to evaluate the performance of candidate grasps by sampling
perturbations in shape, pose, and gripper approach and to compute the probability of force closure for each candidate to
identify a grasp with the highest expected quality. Since evaluating the
quality of each grasp is computationally demanding, prior work has turned to cloud computing. To improve computational efficiency and to extend this work, we consider how
Multi-Armed Bandit (MAB) models for optimizing decisions can be applied in this context. We formulate robust grasp
planning as a MAB problem and evaluate convergence times towards an optimal grasp candidate using 100 object shapes from
the Brown Vision 2D Lab Dataset with 1000 grasp candidates per object. We consider the case where shape uncertainty is represented as a Gaussian process implicit surface (GPIS) with Gaussian uncertainty in pose, gripper approach angle, and coefficient of friction. 
We find that Thompson Sampling and the Gittins index MAB methods converged to within 3\% of the optimal grasp up to 10x faster than uniform allocation and 5x faster than iterative pruning. 
\end{abstract}

@@

     \caption{Number of samples (i.e. iterations of arm pulls) versus the normalized probability of force closure $P_F$ for the best estimated grasp 
        after $t$ samples, $P_F(\Gamma_{\bar{k}, t})$, out of 1000 candidate grasps using uniform allocation, iterative pruning (eliminating candidates that perform poorly on initial samples), and our proposed Multi-Armed Bandit (MAB) algorithms (Gittins indices and Thompson Sampling).
    %We evaluated performance on 100 objects randomly selected from from the Brown Vision 2D Lab Dataset \cite{brown} with 1000 uniformly distributed candidate grasp plans.
    The normalized $P_F$ is given by the ratio of $P_F(\Gamma_{\bar{k}, t})$ to the highest $P_F$ value in the candidate
    grasp set $P_F(\Gamma^*)$ averaged over 100 independent runs on randomly selected objects from the Brown Vision 2D Dataset \cite{brown}.
    The highest quality grasp was determined by brute force search over all candidate grasps (which requires 10x more
    iterations than all methods shown)~\cite{kehoe2012toward}.
    Uniform allocation and iterative pruning converge to within 3$\%$ of the highest quality grasp (the dashed grey line) in approximately 40,000 and 20,000 iterations, respectively. In comparison, the MAB methods both converge in approximately 4,000 iterations. 
%    $P_F(\Gamma_{\bar{k}})$ (ratio of $P_F(\Gamma_{\bar{k}})$ to $P_F(\Gamma^*)$)
%    
%    Normalized grasp quality over time of the returned grasp for 4 sample-based planning methods.  Normalization divides the current quality of grasp, $P_F(\Gamma_{\bar{k}})$ returned by the highest probability of force closure in the candidate grasp set $P_F(\Gamma^*)$. The highest quality grasp was determined by brute force search over all candidate grasps (which required 10x  more time than any of these methods \cite{kehoe2012toward}).  Data is based on 100 objects randomly selected from from the Brown Vision 2D Lab Dataset \cite{brown} with 1000 uniformly distributed candidate grasp plans. Monte Carlo methods such as Uniform sampling and iterative pruning (eliminating candidates that perform poorly on initial samples) eventually converge to within 3$\%$ of the optimal (the dashed grey line), requiring approximately 40,000 and 20,000 iterations.  Two Multi-Arm Bandit methods perform significantly better, converging after only 4000 iterations..
	}
    \label{fig:grasp_quality}
\end{figure}

@@

We compare Thompson sampling and Gittins indices, two popular algorithms for solving the MAB problem,
against uniform allocation and an adaptive sampling method known as iterative pruning, which iteratively reduces the set
of candidate grasps based on the sample mean~\cite{kehoe2012toward},
in terms of the number of samples needed to find a grasp with highest estimated probability of force closure on objects in the Brown Vision 2D Dataset, a dataset of 2D planar objects~\cite{brown, christopoulos2007handling}. Our simulation results show that Thompson Sampling, a MAB algorithm, required $5$x fewer samples than iterative
pruning and $10$x fewer samples than uniform allocation to determine a grasp within $3\%$ of the estimated highest probability of force closure
grasp among a set of 1000 grasp candidates per object and averaged over 100 objects.

@@

One benefit of the Beta prior on Bernoulli reward distributions is that updates to the belief distribution after observing rewards from arm pulls can be derived in closed form.
At timestep $t=0$, we pull arm $k$ and observe reward $R_{k,0}$, where $R_{k,0} \in \lbrace 0, 1\rbrace$.
The posterior of the Beta distribution after this observation is $\alpha_{k, 1} = \alpha_{k, 0} + R_{k,0}$,  $\beta_{k, 1} = \beta_{k, 0} + 1-R_{k,0}$, where $\alpha_{k,0}$ and $\beta_{k,0}$ are the prior shape parameters for arm $k$ before any samples are evaluated.

@@

\subsubsection{Thompson Sampling}
The computational cost of determining the Gittins indices can increase rapidly as the discount factor approaches $1$.
However, in the case of finding the best arm, we want to plan for long-term reward and thus want $\gamma$ as close to $1$ as possible.
Due to computational constraints we must use a smaller $\gamma$, but this can lead to the algorithm pulling only the most promising arm for many iterations~\cite{kelly1981multi}.

@@

To try and reduce the number of samples needed, we instead maximize the sum of $P_F$ values for each sampled grasp  $\Gamma_{k,t}$ at time $t$ up to a given time $T_s$:
\begin{align}
    \underset{\Gamma_{k,*} \in \mG}{\text{max }} \sum \limits_{t=1}^
    {T_s} P_F \left( \Gamma_{k,t}\right)  \label{eq:problem_def}
\end{align}
\noindent The goal is to perform as well as Equation~\ref{eq:old_problem_def} in as few samples as possible~\cite{srinivas2009gaussian}. We then formulate the  problem as a MAB model and compare two different Bayesian MAB algorithms, Thompson sampling and Gittins indices.
